---
title: "Case Study 01 Breweries & Beer"
author: "Hien Lam"
date: "10/8/2021"
output:
  html_document:
    df_print: paged
abstract: 'MSDS6306: Doing Data Science - Case Study 01: Beer & Brewery. I cleaned
  the data and performed EDA to visualize distribution (log transformed when necessary)
  as well as summary statistics. I conducted various hypothesis tests, k-NN cross
  validation on normalized values, linear regression model, and correlation tests
  to derive r, its p-value, r squared. Lastly, I examined the geographical distribution
  of IBU in order to recommend appropriate beer attributes that cater to the unmet
  needs of the regional markets.'
---
## Read & tidy the data

Load the required packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(maps)
library(ggmap)
library(viridis)
library(plotly)
library(caret) #confusion matrix
library(e1071)
library(class) #knn.cv
```
Import the data
```{r message=FALSE, warning=FALSE}
breweries = read_csv('https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv')
beer = read_csv('https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Beers.csv')
```
Breweries data has 558 observations and four columns:

1. Brew_ID: Unique identifier of the brewery  
2. Name: Name of the brewery  
3. City: City where brewery is located  
4. State: U.S. State where brewery is located
```{r}
dim(breweries)
spec(breweries)
```
Beer data has 2410 observations and seven columns:

1. Name: Name of the beer  
2. Beer_ID: Unique identifier of the beer  
3. ABV: Alcohol by volume of the beer  
4. IBU; International Bitterness Units of the beer  
5. Brewery_id: Brewery identifier associated with the beer  
6. Style: Style of the beer  
7. Ounces: Ounces of the beer
```{r}
dim(beer)
spec(beer)
```
Merge datasets by primary key
```{r}
df = merge(breweries, beer, by.x = 'Brew_ID', by.y = 'Brewery_id')
head(df)
```
Rename Name.x and Name.y
```{r}
df = rename(df, Brewery = Name.x, Beer = Name.y)
head(df)
```
Clean the data:

1. Check the sum of NA values in each column (ABV: 62, IBU: 1005, Style: 5)  
2. Create new dataframe excluding the NA values (but also keep original dataframe)  
3. Confirm there are zero NA values left
```{r}
cbind(lapply(lapply(df, is.na), sum))
final = df %>% na.exclude()
```
```{r}
sum(as.numeric(is.na.data.frame(final)))
```

---
## Study the dataset

Final dataset has 1403 observations and 10 columns:  
1. Brew_ID: Unique identifier of the brewery  
2. Brewery: Name of the brewery  
3. City: City where brewery is located  
4. State: U.S. state where brewery is located  
5. Beer: Name of the beer  
6. Beer_ID: Unique identifier of the beer  
7. ABV: Alcohol by volume of the beer  
8. IBU: International Bitterness Units of the beer  
9. Ounces: Ounces of the beer  
10. Style: Style of the beer
```{r}
dim(final)
str(final)
```
The first six and last rows of the dataframe
```{r}
head(final)
tail(final)
```

---
## Exploratory Data Analysis

There are 558 total breweries. I've plotted their distribution by state.
```{r}
breweries %>% group_by(State) %>% tally(sort=T)
```
Plot into a heatmap
```{r}
states = map_data('state')
st_abb = data.frame(abb = state.abb, region = tolower(state.name))
brew = breweries %>% group_by(State) %>% tally(sort=T)
brew = subset(brew, State != 'HI' & State != 'AK') 
brew = rename(brew, abb = State, count = n)
brew = merge(brew, st_abb, by='abb')
geo = merge(states, brew, by='region', all.x=T)
geo = geo[order(geo$order),]
center <- data.frame(region=tolower(state.name), long=state.center$x, lat=state.center$y)
center <- merge(center, brew, by="region", all.x = TRUE)
```
```{r warning=FALSE}
ggplot(geo, aes(x=long,y=lat)) +
geom_polygon(aes(group=group, fill=count)) +
geom_text(data=center, aes(long, lat, label=count)) +
scale_fill_gradient(low = "slategray1",
high = "royalblue4",
guide = "colorbar") +
ggtitle("Breweries per State") +
coord_map() + theme(axis.title = element_text(size = 20)) + theme_void() + theme(legend.position = c(0.9, 0.4), plot.title = element_text(size= 20, hjust=0.1, margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")))
```

I computed median ABV and IBU for each state. Maine has the highest median ABV at 6.7% as well as the highest median IBU at 61.
```{r}
median = final %>% select(State, ABV, IBU) %>% group_by(State) %>% summarize(med_abv = median(ABV), med_ibu = median(IBU))
arrange(median, -med_abv)
arrange(median, -med_ibu)
```

Kentucky has the max ABV at 12.5%
```{r}
max(final$ABV)
grep(.125, final$ABV)
final$State[9]
```

Oregon has the highest IBU at 138
```{r}
max(final$IBU)
grep(138, final$IBU)
final$State[1132]
```
I plotted median ABV for each state.
```{r}
median %>% mutate(perc = med_abv*100, State = fct_reorder(State, perc)) %>% ggplot(aes(State, perc)) + geom_point(size=4, color="royalblue2") + coord_flip() + labs(y='ABV (%)', title='Median Alcohol by Volume of each State', subtitle = 'Decreasing Order') + theme_minimal() + theme(plot.title = element_text(size= 20), axis.title = element_text(size = 20))
```

Median IBU for each state.
```{r}
median %>% mutate(State = fct_reorder(State, med_ibu)) %>% ggplot(aes(State, med_ibu)) + geom_point(size=4, color="indianred2") + coord_flip() + labs(y='IBU', title='Median International Bitterness Unit of each State', subtitle = 'Decreasing Order') + theme_minimal() + theme(plot.title = element_text(size= 20), axis.title = element_text(size = 20))
```


Summary statistics and distribution of ABV

Statistic | Percent 
--------- | ------- 
Min | 2.7  
Med | 5.7  
Mean | 6  
Max | 12.5  
IQR | 1.8  

Right skew distribution suggests lower production at increasing ABV. Consumer preference (and thus sales) drives demand to infer most lean towards lower concentration of ABV. The median value of 5.7% is a better indicator of the distribution because the mean is sensitive to outliers and gets pulled towards the skewed tail. 

Here is the boxplot & density plot of ABV distribution.
```{r}
summary(final$ABV)
fivenum(final$ABV)
IQR(final$ABV)

final %>% mutate(abvperc = ABV*100) %>% ggplot(aes(abvperc)) + geom_boxplot(fill='royalblue', alpha=.3) + labs(x='ABV (%)', title='Distribution of Alcohol by Volume') + theme_cowplot() + theme(plot.title = element_text(size= 20), axis.ticks.y=element_blank(), axis.title.y=element_blank(), axis.text.y=element_blank(), legend.position = 'none', axis.title = element_text(size = 20)) + geom_segment(aes(x = 6, y = -.05, xend = 6, yend = .05, col='red'))
```

```{r}
final %>% mutate(abvperc = ABV*100) %>% ggplot(aes(abvperc)) + geom_density(fill='indianred2', color='indianred3', alpha=.9) + geom_vline(xintercept=5.7, col='cornflowerblue') + geom_vline(xintercept = 6, linetype='dotted', col='paleturquoise3', size=1.1) + labs(x='ABV (%)', title='Distribution of Alcohol by Volume') + theme_minimal_hgrid() + theme(axis.title = element_text(size = 20), plot.title = element_text(size= 20))
```

Relationship between ABV and IBU using linear regression model.

I performed a log transformation of the abv values due to its skewness and conducted a linear regression model. Just from the eye test, the regression line depicts a positive, linear relationship. In fact, with an R2 of .45, it is estimated that 45% of the variation in IBU is explained by abv. Furthermore, there is strong evidence to suggest abv and IBU are linearly correlated (p-val <.00001).The ABV values were log transformed before plotting.
```{r}
final = final %>% mutate(log_abv = log(ABV))
cor.test(x=final$log_abv, y=final$IBU)
model = lm(IBU~log_abv, data=final)
summary(model)
```
ABV vs IBU
```{r}
final %>% ggplot(aes(x=log_abv, y=IBU)) + geom_point(color="black", fill="#69b3a2", shape=22, alpha=0.5, size=3, stroke = 1) + geom_line(aes(x = log_abv,y = model$fit), color = "lightpink3", lwd=.8) + labs(title='Relationship between Bitterness of Beer and its Alcohol Content', subtitle = 'Linear Regression Model on Log Transformed ABV', x= 'ABV (logged)', y='IBU') + theme_classic() + theme(axis.title = element_text(size = 20), plot.title = element_text(size= 20))
```

ABV vs. IBU by Style

View all 90 different styles
```{r}
final %>% group_by(Style) %>% tally(sort=T)
```

Style df of just IPA/ale styles. 32 different styles of IPA and ale encompassing 944 observations and 10 columns
```{r}
style = final %>% group_by(Style) %>% filter(grepl('\\sAle|IPA$', Style))
dim(style)
```
Distinguish ale from ipa
```{r}
all_style = final %>% group_by(Style)

keep = grep('\\sAle', all_style$Style)
ale = all_style[keep,]
rem = grep('India', ale$Style)
ale = ale[-rem,] #552 obs, 27 ale styles
ale$Style = 'ale'

ipa = grep(('\\sIndia\\sPale\\sAle|IPA$'), all_style$Style)
ipa = all_style[ipa,] # 392 obs, 5 ipa styles
ipa$Style = 'ipa'

ipa_ale = rbind(ipa, ale)
ipa_ale$Style = as.factor(ipa_ale$Style)
```
Code the plots
```{r}
pmain = ipa_ale %>% ggplot(aes(x=ABV, y=IBU, shape=factor(Style))) + geom_point(aes(color=factor(Style))) + theme(legend.position = 'none', axis.title = element_text(size = 20), panel.background = element_rect(fill = "gray97", color = NA))

xdens = axis_canvas(pmain, axis='x') + geom_density(data=ipa_ale, aes(ABV, fill=Style), alpha=.7, size=.2)
ydens = axis_canvas(pmain, axis='y', coord_flip = TRUE) + geom_density(data=ipa_ale, aes(IBU, fill=Style), alpha=.7, size=.2) + coord_flip()
p1 = insert_xaxis_grob(pmain, xdens, grid::unit(.2, 'null'), position='top')
p2 = insert_yaxis_grob(p1, ydens, grid::unit(.2, 'null'), position='right')
```
Plot scatterplot with marginal plot on the axis
```{r}
ggdraw(p2) + draw_label('ABV vs. IBU by Style', x=.12, y=.98, size=20) + draw_label('IPA', x=.52, y=.95, size=13, color='paleturquoise4', fontface = 'bold') + draw_label('Ale', x=.35, y=.97, size=13, color='rosybrown', fontface = 'bold')
```

Conduct KNN classifier to investigate the difference with respect to ABV and IBU between IPAs and other types of Ale. I normalized both variables and performed a knn cross validation test with a k of 11. The model achieved a 78% accuracy (how well it correctly identified the style), almost 84% sensitivity (how many ales did the model correctly identify), and 70% specificity (how many ipa did the model correctly identify). Additionally, I log transformed abv and ibu conducted a Welch’s two sample t-test to conclude that there is strong evidence to suggest the true difference in both median abv and ibu value is different for ale and ipa. 
```{r}
normalize = function(x) {
return ((x - min(x)) / (max(x) - min(x)))
                        }
ipa_ale = ipa_ale %>% mutate(norm_abv = normalize(ABV), norm_ibu = normalize(IBU))

model = knn.cv(ipa_ale[,c(12:13)], ipa_ale$Style, k=11)
confusionMatrix(model, ipa_ale$Style)
```
Conduct t.test
```{r}
ipa_ale = ipa_ale %>% mutate(logabv=log(ABV), logibu=log(IBU))
t.test(logabv~Style, data=ipa_ale)
t.test(logibu~Style, data=ipa_ale)
```

---
## INSIGHT

Portland has the highest number of breweries with 39 and its max IBU is 103. Similarly, SD has the third highest number of breweries with 35 and its max IBU is 115. I’ve only given you two examples (due to time constraint), but there is a strong linear correlation between breweries per city and its maximum IBU (correlation coefficient of .79 and p-value of less than .00001). Additionally, it is estimated that 63% of the variation in max IBU is explained by its linear relationship with breweries per city. This leads me to infer cities with a high number of breweries have a more developed palate for crafty beers that have high IBU. 

Why is this important? Boston is ranked 12th in breweries per city yet its maximum IBU is only 45. That’s 50% less than the max IBU compared to cities with similar demographics. In fact, if you take the max IBU of all the cities in MA and extract the median, Boston is still 30 IBU lower. I believe that Boston is an untapped market with a need for higher IBU beer production.

IBU distribution on US map (static plot)
```{r include=FALSE}
register_google('AIzaSyBL5-WnMTC5pWqWzOTfCRmhQoWB2vMpOxQ')
```

```{r message=FALSE}
states = map_data('state')
usabeer = final %>% select(City, State, ABV, IBU)
usabeer = subset(usabeer, State != 'HI' & State != 'AK')
usabeer$citystate = str_c(usabeer$City, ", ", usabeer$State)
usabeer = cbind(geocode(as.character(usabeer$citystate)), usabeer)
usabeer = usabeer %>% arrange(IBU) %>% mutate(cs=factor(citystate, unique(citystate)))

ggplot() + geom_polygon(data=states, aes(x=long, y=lat, group=group), fill='grey', alpha=.3) + geom_point(data=usabeer, aes(lon, y=lat, size=IBU, color=IBU, alpha=IBU), shape=20) + scale_size_continuous(name='IBU', range=c(1,10), breaks=c(1, 25, 50, 75, 100)) + scale_alpha_continuous(name='IBU', range=c(.1, .9), breaks=c(1, 25, 50, 75, 100)) + scale_color_viridis(option='magma', direction=-1, breaks=c(1, 25, 50, 75, 100), name='IBU') + theme_void() + coord_map() + borders('state') + guides( colour = guide_legend()) +
ggtitle("Distribution of International Bitterness Unit by Cities") +
theme(
legend.position = c(0.85, 0.25),
text = element_text(color = "#22211d"),
plot.background = element_rect(fill = "#f5f5f2", color = NA),
panel.background = element_rect(fill = "#f5f5f2", color = NA),
legend.background = element_rect(fill = "#f5f5f2", color = NA),
plot.title = element_text(size= 20, hjust=0.1, margin = margin(b = -.5, t = 0.4, l = 2, unit = "cm")),
)
```
Plot IBU distribution
```{r}
summary(final$IBU)
usabeer %>% ggplot(aes(IBU)) + geom_density(fill='royalblue', alpha=.7) + geom_vline(xintercept=35, col='red') + geom_vline(xintercept = 42.74, linetype='dotted', col='turquoise', size=1.1) + labs(title='International Bitterness Unit') + theme_minimal_hgrid() + theme(axis.title = element_text(size = 20), plot.title = element_text(size= 20))
```

IBU distribution on US map (PLOTLY)
```{r warning=FALSE}
usabeer = usabeer %>% mutate( mytext=paste(citystate, "\n", "IBU: ", IBU, sep=""))

p.usa = ggplot() + geom_polygon(data=states, aes(x=long, y=lat, group=group)) + geom_point(data=usabeer, aes(x=lon, y=lat, size=IBU, color=IBU, alpha=IBU, text=mytext)) + scale_size_continuous(range=c(1,7)) + scale_color_viridis(option='inferno', direction = -1) + theme_void() + coord_map() + borders('state', fill='whitesmoke', alpha=.3) + labs(title='Distribution of International Bitterness Unit by Cities') + theme(legend.position = 'none', plot.title = element_text(size=25, hjust=.5))
p.usa = ggplotly(p.usa, tooltip='text')
p.usa
```

Number of breweries per city in descending order
```{r}
nbrew = final %>% select(IBU, City, State) %>% group_by(City, State) %>% tally(sort=T)
nbrew
```
MAX IBU for each city 
```{r}
maxibu = final %>% select(IBU, City, State) %>% group_by(State, City) %>% slice(which.max(IBU))
maxibu = arrange(maxibu, -IBU)
maxibu
```

merge, conduct cor test
```{r}
ibu.brew.cor = as.data.frame(cbind(maxibu$IBU, nbrew$n))
ibu.brew.cor = rename(ibu.brew.cor, max_ibu=V1, n_brew = V2)
cor(ibu.brew.cor) # .7939
cor.test(x=ibu.brew.cor$n_brew, y=ibu.brew.cor$max_ibu, method = 'pearson')

```

```{r}
ma = maxibu %>% select(State, IBU) %>% filter(State == 'MA') %>% summarize(value = IBU)
mean(ma$value)
fivenum(ma$value)
```



